---
title: "Eflows Species Clustering in CA Regions"
output:
  html_document:
    df_print: paged
    css: mystyle.css
---

```{r setup, echo=FALSE, include=FALSE}
library(sf)
library(maps)
#library(devtools)
#devtools::install_github("tidyverse/ggplot2") # need latest version of ggplot2 in order to use geom_sf
library(ggplot2)
library(rgdal)
library(tidyr)
library(dplyr)

group_palette <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a')

default_num_groups <- c(3,4,5,6,7)

huc12_eflow_percents <- read.csv("data/HUC12s_eflow_percent.csv", stringsAsFactors = FALSE)
suffix<-"groups_num"

huc_data <- sf::st_as_sf(rgdal::readOGR(paste("data/report_update/clusters2.gpkg", sep=""),"june_2018_current_notrans_neighbors", stringsAsFactors=FALSE))

filter_data_to_group <- function (group_name, data){
  return(data %>% drop_na(paste(group_name,'_5groups_num', sep="")))    
}

species_maps <-function(groupname, title, region_data){
    
      huc_data <- filter_data_to_group(groupname, region_data)  
  
      # now, we need to convert the attributes we're using for groupings into a single column so that GGPlot can do our mapping for us with facets
      # paste(groupname,"_2",suffix, sep=""),paste(groupname,"_8",suffix, sep=""),paste(groupname,"_9",suffix, sep="")
      huc_data_mapping <- huc_data %>% gather(key=model_run, value=grouping, paste(groupname,"_3",suffix, sep=""),paste(groupname,"_4",suffix, sep=""),paste(groupname,"_5",suffix, sep=""),paste(groupname,"_6",suffix, sep=""),paste(groupname,"_7",suffix, sep=""))
      
      plot <- ggplot() + 
        geom_sf(huc_data_mapping, lwd = 5, mapping=aes(fill=factor(grouping)), color=NA)+
        scale_fill_manual(values=group_palette, limits=c(1,2,3,4,5,6,7,8,9,10)) +
        ggtitle(paste(title," Species Groupings", sep="")) +
        facet_wrap("model_run")+
        theme_bw()+
        theme(panel.grid.major = element_line(colour = 'transparent'))+
        theme(strip.background = element_blank(), # facet theme control
              strip.placement = "outside", 
              axis.line.x = element_line(color="black", size = 1),
              axis.line.y = element_line(color="black", size = 1),
              plot.title = element_text(hjust = 0.5))
  return(list("region_map"=plot,"species_data"=huc_data))
      
}


species_lists_by_group_dfs <- function (data_frame, group_name, groups=default_num_groups, counts=FALSE){

  data_frame <- filter_data_to_group(group_name, data_frame)
  st_geometry(data_frame) <- NULL # drop the SF geometry info
  remove_fields <- c("OBJECTID", "huc_8", "huc_10", "huc_12_string", "huc_12_double", "taxa_count", "geom_Length", "geom_Area", "oid_int_copy", "Shape_Length", "Shape_Area", "huc4","huc_region_group")
  
  data_of_interest <- select(data_frame, -one_of(remove_fields))  # filters to just the current group and removes extra fields
  
  current_field_names <-names(data_of_interest) # get the current column names
  new_field_names <- gsub("Flow_Sensitive_current_", "", current_field_names)  # replace the prefix on the fields
  
  colnames(data_of_interest) <- new_field_names
  
  # general approach we'd want here is to make vectors of the species in each group, then add to a list
  # then extend all lists to be the same length, then cbind them all to a single df and print the df
  all_groupings = list()
  for(grouping in groups){
    output_list = list()  # make an empty output list
    field_name <- paste(group_name,"_",grouping,suffix, sep="")
    for(group in seq(1,grouping)){ # for each subgroup in this grouping
      records_in_group <- data_of_interest %>% dplyr::filter_(paste(field_name, "==", as.character(group)))  # filters to just the current group

      records_in_group <- records_in_group[, -grep(".+groups_.+", colnames(records_in_group))]  # drop columns for groupings
      gathered_data <- as.data.frame(records_in_group) %>% tidyr::gather(key="species", value="presence")
      found_species <- gathered_data %>% dplyr::filter(presence==1) %>% distinct(species)
      g_title <- paste("group_", group, sep="")
      #colnames(found_species) <- g_title  # assign it so that the data frame field has the right name before we merge it
      #found_species <- sort(found_species, na.last=TRUE)
      output_list[[g_title]] = found_species$species  # add the group to the output list  
      
    }
    output_list <- lapply(output_list, `length<-`, max(lengths(output_list)))  # Make all of the vectors in the list the same length
    
    all_groupings[[grouping]] <- as.data.frame(output_list)
  }
  
  return(groupings=all_groupings)
}


get_dominance_data  <- function(data_frame, group_name, groups=default_num_groups, counts=FALSE){
  st_geometry(data_frame) <- NULL # drop the SF geometry info
  group_dominance_data = data.frame(number_clusters=numeric(), value=numeric(), dominant_class=character())  # set it to NA to start - a function will create a data frame from it
  for(grouping in groups){
    field_name <- paste(group_name,"_",grouping,suffix, sep="")
    group_dominance_data <- get_cluster_dominance_data(data_frame, field_name, grouping, group_dominance_data, counts=counts)  # pass in the whole DF with the extra fields - we need the HUC 12 field
  }
  return(group_dominance_data)
}

get_dominant_stream_class_by_hucs <- function(group_records, eflow_data=huc12_eflow_percents, counts=FALSE){
  ## Given a set of records from a cluster, determines the dominant stream flow class in that cluster
  ## and the proportion of it that is the dominant class
  ## If counts is TRUE, then instead of proportion, it outputs the count as the "proportion_dominant", but
  ## it will still include the dominant class in the list output for symbolization
  
  eflow_records <- eflow_data %>% filter(HUC12 %in% group_records$huc_12_string)
  eflow_sums <- eflow_records %>% group_by(eflow_type) %>% summarise(group_sum = sum(eflow_length, na.rm=TRUE))
  max_record = which.max(eflow_sums$group_sum)
  proportion_dominant = eflow_sums[max_record,]$group_sum / sum(eflow_sums$group_sum)
  if(counts==TRUE){
    output_value = nrow(eflow_sums)
  }else{
    output_value = proportion_dominant
  }
  return(list(class=eflow_sums[max_record,]$eflow_type, result_value=output_value))
}

add_record_to_df <- function(original_df, num_groups, percent, dominant_class){
  ## This record will keep adding data to the end of the data frame about the dominance data
  if(length(percent) == 0 || length(dominant_class) == 0){  # in some cases, we get an effective null, but it's numeric(0) - we have to test the length of it to get whether or not it's this value
    return(original_df)
  }
  new_group_data = data.frame(number_clusters=num_groups, value=percent, dominant_class=dominant_class)
  
  return(rbind(original_df, new_group_data))  # otherwise, append the new data to the end
}

get_cluster_dominance_data <- function(records, cluster_field, num_clusters, existing_data, counts=FALSE){
  # make an empty data frame to hold the data
  for(group_num in 1:num_clusters){
    group_records <- records %>% dplyr::filter_(paste(cluster_field, "==", group_num))
    group_info <- get_dominant_stream_class_by_hucs(group_records, counts=counts)
    existing_data <- add_record_to_df(existing_data, num_clusters, group_info$result_value, group_info$class)
  }
  
  return(existing_data)

}

plot_dominance_gg <- function(dominance_data, title){
  
  # coloration via: https://stackoverflow.com/a/21146026/587938
  new_plot <- ggplot(dominance_data, aes(x=number_clusters, y=value)) +
              geom_point(mapping=aes(colour=dominant_class, size=5)) + 
              scale_colour_discrete(drop=TRUE, limits = levels(dominance_data$dominant_class)) +
              geom_smooth(method = "lm") +
              ggtitle(title)
  
  return(new_plot)
}

print_dominance_plots <- function(species_info, region_name){
  dominance_data <- get_dominance_data(species_info$species_data, region_name, counts=FALSE)
  print(plot_dominance_gg(dominance_data, title=paste("Percent Dominance by K Value in", region_name)))
  dominance_data <- get_dominance_data(species_info$species_data, region_name, counts=TRUE)
  print(plot_dominance_gg(dominance_data, title=paste("Number of Stream Classes by K Value in", region_name)))
}


```


## Overview and Basic Methods
In this version (June 2018) of the clustering, translocations were _not_ included. Clusters must be contiguous in this version, as determined by a shared boundary or vertex, not by hydrologic connectivity. All HUCs were included, regardless of the minimum number of taxa (includes ones with one or none). Clustering was done with Flow Sensitive Taxa.

Clustering was done by a PISCES tool that subsets and formulates the data to be used by ArcGIS' Grouping Analysis tool, where prior versions were clustered through manual steps and then fed into GeoDa's K-Means tool.


## Map of Region Breakouts
The region was broken out primarily into
 * North Coast (North Coast->San Francisco)
 * South Coast (South of SF -> Southern California)
 * Central Valley (anything draining out through the Central Valley - excludes Honey/Eagle Lakes and SF Bay)
 * Great Basin (anything in hydrologic region 16 in the northern part of the state, contiguous)
 * Desert - not clustered or shown here as previous clustering indicated single-species driven areas (Includes eastern sierra that is part of region 18, souther reaches of region 15 in the state, portion of Colorado River, and a small piece of region 16)

The map below shows the areas that were actually included in clustering - in this case, that also means that areas with less than two taxa were removed.

```{r map of region breakouts, fig.height=12, fig.width=12,}

writeLines("td, th { padding : 6px } th { background-color : #fff ; color : #000; border : 1px solid #333; } td { color : #000 ; border : 1px solid #333 }", con = "mystyle.css")  # add the CSS for the data tables

ggplot() + 
  geom_sf(huc_data, lwd = 5, mapping=aes(fill=huc_region_group), color=NA)+ 
  ggtitle("Region Breakout") + 
  theme_bw()+
  theme(panel.grid.major = element_line(colour = 'transparent'))+
  theme(strip.background = element_blank(), # facet theme control
        strip.placement = "outside", 
        axis.line.x = element_line(color="black", size = 1),
        axis.line.y = element_line(color="black", size = 1),
        plot.title = element_text(hjust = 0.5))


```

## Central Valley

```{r map central valley, fig.height=12, fig.width=12, results='asis'}
cv_region_name = "central_valley"
cv_species_info <- species_maps(cv_region_name, "Central Valley", huc_data)
cv_species_info$region_map  # print the map
print_dominance_plots(cv_species_info, cv_region_name)
cv_species_lists <- species_lists_by_group_dfs(cv_species_info$species_data, cv_region_name)  # species_info$species_data is the data frame for the region

for(df in cv_species_lists){
  show(knitr::kable(df, format="html"))
}



```


## Great Basin

```{r map great basin, fig.height=12, fig.width=12, results='asis'}
gb_region_name <- "great_basin"
gb_species_info <- species_maps(gb_region_name, "Great Basin", huc_data)
gb_species_info$region_map  # print the map
print_dominance_plots(gb_species_info, gb_region_name)
gb_species_lists <- species_lists_by_group_dfs(gb_species_info$species_data, gb_region_name)  # species_info$species_data is the data frame for the region

for(df in gb_species_lists){
  show(knitr::kable(df, format="html"))
}

```


## North Coast

```{r map coast, fig.height=12, fig.width=12, results='asis'}
coast_region_name = "north_coast"
coast_species_info <- species_maps(coast_region_name, "North Coast", huc_data)
coast_species_info$region_map  # print the map
print_dominance_plots(coast_species_info, coast_region_name)
coast_species_lists <- species_lists_by_group_dfs(coast_species_info$species_data, coast_region_name)  # species_info$species_data is the data frame for the region

for(df in coast_species_lists){
  show(knitr::kable(df, format="html"))
}

```

## South Coast

```{r map South Coast, fig.height=12, fig.width=12, results='asis'}
soco_region_name = "south_coast"
soco_species_info <- species_maps(soco_region_name, "South Coast", huc_data)
soco_species_info$region_map  # print the map
print_dominance_plots(soco_species_info, soco_region_name)
soco_species_lists <- species_lists_by_group_dfs(soco_species_info$species_data, soco_region_name)  # species_info$species_data is the data frame for the region

for(df in soco_species_lists){
  show(knitr::kable(df, format="html"))
}

```



## For meeting on k groups
```{r meeting figures, fig.height=6, fig.width=6, results='asis', echo=FALSE, eval=FALSE}
## NOTE THE EVAL=FALSE - this chunk won't run when knit - just for interactive work in meeting!
# make coast info
coast_region_name = "coast"
coast_species_info <- species_maps(coast_region_name, "Coast", huc_data)
coast_species_info$region_map  # print the map
group_data <- species_lists_by_group_dfs(coast_species_info$species_data, coast_region_name, counts=FALSE)  # species_info$species_data is the data frame for the region
dominance_data <- group_data$dominance_data
print(plot_dominance_gg(dominance_data))
group_data <- species_lists_by_group_dfs(coast_species_info$species_data, coast_region_name, counts=TRUE)  # species_info$species_data is the data frame for the region
dominance_data <- group_data$dominance_data
print(plot_dominance_gg(dominance_data))

# make CV
cv_region_name = "centralvalley"
cv_species_info <- species_maps(cv_region_name, "Central Valley")
cv_species_info$region_map  # print the map
cv_group_data <- species_lists_by_group_dfs(cv_species_info$species_data, cv_region_name)  # species_info$species_data is the data frame for the region
cv_dominance_data <- cv_group_data$dominance_data

print(plot_dominance_gg(cv_dominance_data))

```

