---
title: "Eflows Species Clustering in CA Regions"
output:
  html_document:
    df_print: paged
---

```{r setup, echo=FALSE, include=FALSE}
library(sf)
library(maps)
#library(devtools)
#devtools::install_github("tidyverse/ggplot2") # need latest version of ggplot2 in order to use geom_sf
library(ggplot2)
library(rgdal)
library(tidyr)
library(dplyr)

group_palette <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a')

species_maps <-function(groupname, title){
  suffix<-"_R50_Std_AM_Euc"
        huc_data <- sf::st_as_sf(rgdal::readOGR(paste("huc_groupings/",groupname,"_flowsensitive.gpkg", sep=""),groupname, stringsAsFactors=FALSE))
      # now, we need to convert the attributes we're using for groupings into a single column so that GGPlot can do our mapping for us with facets
      huc_data_mapping <- huc_data %>% gather(key=model_run, value=grouping, paste(groupname,"_KM2",suffix, sep=""),paste(groupname,"_KM3",suffix, sep=""),paste(groupname,"_KM4",suffix, sep=""),paste(groupname,"_KM5",suffix, sep=""),paste(groupname,"_KM6",suffix, sep=""),paste(groupname,"_KM7",suffix, sep=""),paste(groupname,"_KM8",suffix, sep=""),paste(groupname,"_KM9",suffix, sep=""))
      
      plot <- ggplot() + 
        geom_sf(huc_data_mapping, lwd = 5, mapping=aes(fill=grouping), color=NA)+ 
        scale_fill_manual(values=group_palette, limits=c(1,2,3,4,5,6,7,8,9)) +
        ggtitle(paste(title," Species Groupings", sep="")) +
        facet_wrap("model_run")+
        theme_bw()+
        theme(panel.grid.major = element_line(colour = 'transparent'))+
        theme(strip.background = element_blank(), # facet theme control
              strip.placement = "outside", 
              axis.line.x = element_line(color="black", size = 1),
              axis.line.y = element_line(color="black", size = 1),
              plot.title = element_text(hjust = 0.5))
  return(list("region_map"=plot,"species_data"=huc_data))
      
}

species_lists_by_group <- function (data_frame, group_name, groups=c(2,3,4,5,6,7,8,9)){
    st_geometry(data_frame) <- NULL # drop the SF geometry info
  
  suffix<-"_R50_Std_AM_Euc"
  remove_fields <- c("OBJECTID", "hucs_filtered_HUC_8", "hucs_filtered_HUC_10", "hucs_filtered_HUC_12", "hucs_filtered_ACRES", "hucs_filtered_NCONTRB_A", "hucs_filtered_HU_10_GNIS", "hucs_filtered_HU_12_GNIS", "hucs_filtered_HU_10_DS", "hucs_filtered_HU_10_NAME", "hucs_filtered_HU_10_MOD", "hucs_filtered_HU_10_TYPE", "hucs_filtered_HU_12_DS", "hucs_filtered_HU_12_NAME", "hucs_filtered_HU_12_MOD", "hucs_filtered_HU_12_TYPE", "hucs_filtered_STATES", "hucs_filtered_CVA_NAME", "hucs_filtered_huc_id", "Flow_Sensitive_current_OBJECTID", "Flow_Sensitive_current_HUC_12", "hucs_filtered_META_ID", "Shape_Length", "Shape_Area", "huc4","huc_region_group",paste("KM_6",suffix, sep=""),paste("KM_5",suffix, sep=""),paste("KM_8",suffix, sep=""),paste("KM_10",suffix, sep=""),paste("KM_12",suffix, sep=""),paste("KM_14",suffix, sep=""))
  
  data_of_interest <- select(data_frame, -one_of(remove_fields))  # filters to just the current group and removes extra fields
  
  current_field_names <-names(data_of_interest) # get the current column names
  new_field_names <- gsub("Flow_Sensitive_current_", "", current_field_names)  # replace the prefix on the fields
  
  colnames(data_of_interest) <- new_field_names
  
  for(grouping in groups){
    writeLines(paste("\n\n",grouping, "Groups", sep=" "))
    for(group in seq(1,grouping)){ # for each subgroup in this grouping
      writeLines(paste("\nGroup", group, sep=" "))
      
      records_in_group <- data_of_interest %>% dplyr::filter_(paste(paste(group_name,"_KM",grouping,suffix, sep=""), "==", as.character(group)))  # filters to just the current group

      records_in_group <- records_in_group[, -grep(".+KM.+", colnames(records_in_group))]  # drop columns for groupings
      gathered_data <- as.data.frame(records_in_group) %>% tidyr::gather(key="species", value="presence")
      found_species <- gathered_data %>% dplyr::filter(presence==1) %>% distinct(species)
      writeLines(paste(nrow(found_species), "species in group"))
      writeLines(found_species$species)
    }
  }
}

```


## Map of Region Breakouts
The region was broken out primarily into
* Coast (North Coast->Santa Barbara)
* Central Valley (anything draining out through the Central Valley - excludes Honey/Eagle Lakes)
* Great Basin (anything in hydrologic region 16 in the northern part of the state, contiguous)
* Desert (southern coast, including LA, and anything in the desert regions. Includes eastern sierra that is part of region 18, souther reaches of region 15 in the state, portion of Colorado River, and a small piece of region 16)

```{r map of region breakouts, fig.height=12, fig.width=12,}
region_data <- sf::st_as_sf(rgdal::readOGR("groupings/flowsensitive_current.gpkg", "flowsensitive_current", stringsAsFactors=FALSE))

ggplot() + 
  geom_sf(region_data, lwd = 5, mapping=aes(fill=huc_region_group), color=NA)+ 
  ggtitle("Region Breakout") + 
  theme_bw()+
  theme(panel.grid.major = element_line(colour = 'transparent'))+
  theme(strip.background = element_blank(), # facet theme control
        strip.placement = "outside", 
        axis.line.x = element_line(color="black", size = 1),
        axis.line.y = element_line(color="black", size = 1),
        plot.title = element_text(hjust = 0.5))


```

## Central Valley

```{r map central valley, fig.height=24, fig.width=24,}
region_name = "centralvalley"
species_info <- species_maps(region_name, "Central Valley")
species_info$region_map  # print the map
species_lists_by_group(species_info$species_data, region_name)  # species_info[1] is the data frame for the region



```


## Great Basin

```{r map great basin, fig.height=24, fig.width=24,}
region_name = "greatbasin"
species_info <- species_maps(region_name, "Great Basin")
species_info$region_map  # print the map
species_lists_by_group(species_info$species_data, region_name)  # species_info$species_data is the data frame for the region

```


## Coast

```{r map coast, fig.height=24, fig.width=24,}
region_name = "coast"
species_info <- species_maps(region_name, "Coast")
species_info$region_map  # print the map
species_lists_by_group(species_info$species_data, region_name)  # species_info$species_data is the data frame for the region

```

## Desert

```{r map Desert, fig.height=24, fig.width=24,}
region_name = "desert"
species_info <- species_maps(region_name, "Desert")
species_info$region_map  # print the map
species_lists_by_group(species_info$species_data, region_name)  # species_info$species_data is the data frame for the region

```