---
title: "Eflows Species Clustering in CA Regions"
output:
  html_document:
    df_print: paged
    css: mystyle.css
---

```{r setup, echo=FALSE, include=FALSE}
library(sf)
library(maps)
#library(devtools)
#devtools::install_github("tidyverse/ggplot2") # need latest version of ggplot2 in order to use geom_sf
library(ggplot2)
library(rgdal)
library(tidyr)
library(dplyr)

group_palette <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a')

species_maps <-function(groupname, title){
  suffix<-"_R50_Std_AM_Euc"
        huc_data <- sf::st_as_sf(rgdal::readOGR(paste("data/clustering/huc_groupings/",groupname,"_flowsensitive.gpkg", sep=""),groupname, stringsAsFactors=FALSE))
      # now, we need to convert the attributes we're using for groupings into a single column so that GGPlot can do our mapping for us with facets
      huc_data_mapping <- huc_data %>% gather(key=model_run, value=grouping, paste(groupname,"_KM2",suffix, sep=""),paste(groupname,"_KM3",suffix, sep=""),paste(groupname,"_KM4",suffix, sep=""),paste(groupname,"_KM5",suffix, sep=""),paste(groupname,"_KM6",suffix, sep=""),paste(groupname,"_KM7",suffix, sep=""),paste(groupname,"_KM8",suffix, sep=""),paste(groupname,"_KM9",suffix, sep=""))
      
      plot <- ggplot() + 
        geom_sf(huc_data_mapping, lwd = 5, mapping=aes(fill=grouping), color=NA)+ 
        scale_fill_manual(values=group_palette, limits=c(1,2,3,4,5,6,7,8,9)) +
        ggtitle(paste(title," Species Groupings", sep="")) +
        facet_wrap("model_run")+
        theme_bw()+
        theme(panel.grid.major = element_line(colour = 'transparent'))+
        theme(strip.background = element_blank(), # facet theme control
              strip.placement = "outside", 
              axis.line.x = element_line(color="black", size = 1),
              axis.line.y = element_line(color="black", size = 1),
              plot.title = element_text(hjust = 0.5))
  return(list("region_map"=plot,"species_data"=huc_data))
      
}


species_lists_by_group_dfs <- function (data_frame, group_name, groups=c(2,3,4,5,6,7,8,9)){
  # data.frame version of the code
  
    st_geometry(data_frame) <- NULL # drop the SF geometry info
  
  suffix<-"_R50_Std_AM_Euc"
  remove_fields <- c("OBJECTID", "hucs_filtered_HUC_8", "hucs_filtered_HUC_10", "hucs_filtered_HUC_12", "hucs_filtered_ACRES", "hucs_filtered_NCONTRB_A", "hucs_filtered_HU_10_GNIS", "hucs_filtered_HU_12_GNIS", "hucs_filtered_HU_10_DS", "hucs_filtered_HU_10_NAME", "hucs_filtered_HU_10_MOD", "hucs_filtered_HU_10_TYPE", "hucs_filtered_HU_12_DS", "hucs_filtered_HU_12_NAME", "hucs_filtered_HU_12_MOD", "hucs_filtered_HU_12_TYPE", "hucs_filtered_STATES", "hucs_filtered_CVA_NAME", "hucs_filtered_huc_id", "Flow_Sensitive_current_OBJECTID", "Flow_Sensitive_current_HUC_12", "hucs_filtered_META_ID", "Shape_Length", "Shape_Area", "huc4","huc_region_group",paste("KM_6",suffix, sep=""),paste("KM_5",suffix, sep=""),paste("KM_8",suffix, sep=""),paste("KM_10",suffix, sep=""),paste("KM_12",suffix, sep=""),paste("KM_14",suffix, sep=""))
  
  data_of_interest <- select(data_frame, -one_of(remove_fields))  # filters to just the current group and removes extra fields
  
  current_field_names <-names(data_of_interest) # get the current column names
  new_field_names <- gsub("Flow_Sensitive_current_", "", current_field_names)  # replace the prefix on the fields
  
  colnames(data_of_interest) <- new_field_names
  
  # general approach we'd want here is to make vectors of the species in each group, then add to a list
  # then extend all lists to be the same length, then cbind them all to a single df and print the df
  all_groupings = list()
  for(grouping in groups){
    output_list = list()  # make an empty output list
    for(group in seq(1,grouping)){ # for each subgroup in this grouping
      records_in_group <- data_of_interest %>% dplyr::filter_(paste(paste(group_name,"_KM",grouping,suffix, sep=""), "==", as.character(group)))  # filters to just the current group

      records_in_group <- records_in_group[, -grep(".+KM.+", colnames(records_in_group))]  # drop columns for groupings
      gathered_data <- as.data.frame(records_in_group) %>% tidyr::gather(key="species", value="presence")
      found_species <- gathered_data %>% dplyr::filter(presence==1) %>% distinct(species)
      g_title <- paste("group_", group, sep="")
      #colnames(found_species) <- g_title  # assign it so that the data frame field has the right name before we merge it
      #found_species <- sort(found_species, na.last=TRUE)
      output_list[[g_title]] = found_species$species  # add the group to the output list  
    
    }
    output_list <- lapply(output_list, `length<-`, max(lengths(output_list)))  # Make all of the vectors in the list the same length
    
    all_groupings[[grouping]] <- as.data.frame(output_list)
  }
  
  return(all_groupings)
}

```


## Map of Region Breakouts
The region was broken out primarily into
* Coast (North Coast->Santa Barbara)
* Central Valley (anything draining out through the Central Valley - excludes Honey/Eagle Lakes)
* Great Basin (anything in hydrologic region 16 in the northern part of the state, contiguous)
* Desert (southern coast, including LA, and anything in the desert regions. Includes eastern sierra that is part of region 18, souther reaches of region 15 in the state, portion of Colorado River, and a small piece of region 16)

```{r map of region breakouts, fig.height=12, fig.width=12,}

writeLines("td, th { padding : 6px } th { background-color : #fff ; color : #000; border : 1px solid #333; } td { color : #000 ; border : 1px solid #333 }", con = "mystyle.css")  # add the CSS for the data tables

flow_sense_data <- file.path("C:/Users/dsx/Code/eflow-species/data/clustering/groupings/flowsensitive_current.gpkg")
region_data <- sf::st_as_sf(rgdal::readOGR(flow_sense_data, "flowsensitive_current", stringsAsFactors=FALSE))

ggplot() + 
  geom_sf(region_data, lwd = 5, mapping=aes(fill=huc_region_group), color=NA)+ 
  ggtitle("Region Breakout") + 
  theme_bw()+
  theme(panel.grid.major = element_line(colour = 'transparent'))+
  theme(strip.background = element_blank(), # facet theme control
        strip.placement = "outside", 
        axis.line.x = element_line(color="black", size = 1),
        axis.line.y = element_line(color="black", size = 1),
        plot.title = element_text(hjust = 0.5))


```

## Central Valley

```{r map central valley, fig.height=24, fig.width=24, results='asis'}
region_name = "centralvalley"
species_info <- species_maps(region_name, "Central Valley")
species_info$region_map  # print the map
species_lists <- species_lists_by_group_dfs(species_info$species_data, region_name)  # species_info$species_data is the data frame for the region

for(df in species_lists){
  show(knitr::kable(df, format="html"))
}



```


## Great Basin

```{r map great basin, fig.height=24, fig.width=24, results='asis'}
region_name = "greatbasin"
species_info <- species_maps(region_name, "Great Basin")
species_info$region_map  # print the map
species_lists <- species_lists_by_group_dfs(species_info$species_data, region_name)  # species_info$species_data is the data frame for the region

for(df in species_lists){
  show(knitr::kable(df, format="html"))
}

```


## Coast

```{r map coast, fig.height=24, fig.width=24, results='asis'}
region_name = "coast"
species_info <- species_maps(region_name, "Coast")
species_info$region_map  # print the map
species_lists <- species_lists_by_group_dfs(species_info$species_data, region_name)  # species_info$species_data is the data frame for the region

for(df in species_lists){
  show(knitr::kable(df, format="html"))
}

```

## Desert

```{r map Desert, fig.height=24, fig.width=24, results='asis'}
region_name = "desert"
species_info <- species_maps(region_name, "Desert")
species_info$region_map  # print the map
species_lists <- species_lists_by_group_dfs(species_info$species_data, region_name)  # species_info$species_data is the data frame for the region

for(df in species_lists){
  show(knitr::kable(df, format="html"))
}

```