---
title: "MCA Species Regions"
author: "Alyssa Obester"
date: "September 13, 2018"
output: html_document
self_contained: no
---

```{r setup, include=FALSE}
#This article used to generate code:http://www.gastonsanchez.com/visually-enforced/how-to/2012/10/13/MCA-in-R/
#loading all required packages
library(FactoMineR)
library(tidyverse)
library(ggthemes)
library(ggrepel)
library(vegan)

knitr::opts_chunk$set(echo = TRUE)
huc_data <- sf::st_as_sf(rgdal::readOGR(paste("~/GitHub/eflow-species/data/phase2/cluster_family_level.gpkg", sep=""),"family_flowsens_3to8_notrans_nodelta", stringsAsFactors=FALSE))

```

## Dissimilarity Analysis for Regions Used in Clustering
We performed a multiple component analysis for the four geographic regions in California used for clustering: Central Valley, Great Basin, North Coast and South Coast, to evaluate differences in fish species presence within the four regions.

## Map of Regions 
The regions were broken out into:  * North Coast (North Coast->San Francisco)  * South Coast (South of SF -> Southern California)  * Central Valley (anything draining out through the Central Valley - excludes Honey/Eagle Lakes and SF Bay)  * Great Basin (anything in hydrologic region 16 in the northern part of the state, contiguous)  * Desert - not clustered or shown here as previous clustering indicated single-species driven areas (Includes eastern sierra that is part of region 18, souther reaches of region 15 in the state, portion of Colorado River, and a small piece of region 16). See **eflow-species\huc_region_groups.Rmd** for more information on geographic boundaries of the regions. 


```{r, echo = FALSE, fig.height=12, fig.width=12,}

writeLines("td, th { padding : 6px } th { background-color : #fff ; color : #000; border : 1px solid #333; } td { color : #000 ; border : 1px solid #333 }", con = "mystyle.css")  # add the CSS for the data tables

region_boundaries <- 
ggplot() + 
  geom_sf(huc_data, lwd = 5, mapping=aes(fill=huc_region_group), color=NA)+ 
  ggtitle("Region Breakout") + 
  theme_bw()+
  theme(panel.grid.major = element_line(colour = 'transparent'))+
  theme(strip.background = element_blank(), # facet theme control
        strip.placement = "outside", 
        axis.line.x = element_line(color="black", size = 1),
        axis.line.y = element_line(color="black", size = 1),
        plot.title = element_text(hjust = 0.5))

region_boundaries
```


```{r, include = FALSE}
#reading in the data
common_name_regions <- read.csv("~/GitHub/eflow-species/scripts/mca/common_name_regions.csv")
cn_reg <- common_name_regions

head(cn_reg)
str(cn_reg)

# try transforming:
cn_reg_long <- gather(cn_reg, key = "group", value = "species") %>% 
  filter(!is.na(species), !species=="")

# view number of categories per group
(cats = apply(cn_reg_long, 2, function(x) nlevels(as.factor(x))))


```

## Do MCA
```{r}
#run the MCA
mca1 = MCA(cn_reg_long, graph = F)
mca1

#Generates a table of eigenvalues
mca1$eig

#Create a data frame with variable coordinates
mca1_vars_df = data.frame(mca1$var$coord, Variable = rep(names(cats), cats))

#Create a data frame with observation coordinates
mca1_obs_df = data.frame(mca1$ind$coord)
```


### Plot of Variable Categories
``` {r pressure, echo = FALSE}
ggplot(data=mca1_vars_df, 
       aes(x = Dim.1, y = Dim.2, label = rownames(mca1_vars_df))) +
  geom_point(size=1, alpha = 0.7) +
  geom_hline(yintercept = 0, colour = "gray70") +
  geom_vline(xintercept = 0, colour = "gray70") +
  geom_text_repel(aes(colour=Variable)) +
  #scale_colour_colorblind()+
  scale_x_continuous(breaks=seq(-5,5), limits=c(-5,5)) +
  scale_y_continuous(breaks=seq(-5,5), limits=c(-5,5))+
  #geom_density2d(colour = "gray80") +
  ggtitle("MCA plot of variables using R package FactoMineR") 
```

### Make Matrix and Create Dendogram

```{r}
# make a 1/0 matrix 
cn_reg_mat <- cn_reg_long %>% 
  mutate(g1=if_else(group=="central_valley", 1, 0),
         g2=if_else(group=="great_basin", 1, 0),
         g3=if_else(group=="north_coast", 1, 0),
         g4=if_else(group=="south_coast", 1, 0)) %>% 
  select(-group) %>% group_by(species) %>% 
  summarize_all(sum) %>% 
  mutate(species=as.factor(species)) %>% as.data.frame()

cats = apply(cn_reg_mat, 2, function(x) nlevels(as.factor(x)))
cats

# use jaccard for 1/0 (presence absence)
spgrp.mat<-as.matrix(cn_reg_mat[,2:5])
spgrp.dist<-vegdist(spgrp.mat, method = "jaccard")

# plot a heirarchical clustering of the dissimilarity matrix (need to zoom on this plot to make it legible)
plot(hclust(spgrp.dist), labels = cn_reg_mat$species, cex=0.7)
```

